### 2022-10-19 manuscript/manuscript.tex
so tired of theory

\begin{align}
  \label{eq:qterm}
  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i \\
  Q_\infty &= \frac{1}{1-(1-r) \kappa}
\end{align}

which shows the long reach of heritable fitness factors in altering allele
frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
built associations between the neutral allele and a fitness background from
$S_1$. As long as the direction of selection is the same, the expected change
of this fraction of associations formed in the first generation to later
generations 2 and 3 would be,

@@

First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
multiplicative fitness model is assumed (i.e. independent effects, no
epistasis), where the fitness of individual $i$ is the product of their fitness
contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^L
w_{i,l}$. Then, assuming mean fitness is one, the total fitness variation is
multiplicative across the locus-level fitnesses,

\begin{align}
  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^L (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^L (1 + V_{h,l})  - 1
\end{align}

Now, the $Q_t^2 V_h$ term in the single-locus model is now under a
multiplicative polygenic model,

\begin{align}
  Q_t^2 V_h &= \prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}

@@

\begin{align}
  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
          &\approx \frac{N}{\prod_{l=1}^L \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
          &\approx N \exp\left(- \sum_{l=1}^L \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}

(c.f. \cite{Santiago1998-bs} equation 4b). Santiago and Caballero ignore the
covariances between the locus-specific variation decay rates $1-\kappa_l$ terms
and the locus-specific variances, $V_{h,l}$. This allows for the terms to be
separated,

\begin{align}
  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
\end{align}

@@

Our model differs from that of \textcite{Santiago1998-bs} in that we do not
integrate over the entire genome. In their 1998 model, Santiago and Caballero
consider the impact of both sites under selection on the same chromosome, as
well as sites on independently assorting chromosomes. We only consider the
contribution of a single segment under purifying selection, and the reduction
from additional segments accumulate multiplicatively (similar to our
implementation of the classic BGS model, XXX).

If we assume that the $L$ selected sites are evenly distributed in a region of
length $G$ basepairs, we can rewrite the sum as $\nicefrac{L}{G} \sum_{l=1}^G
Q_{t,l}^2$

\begin{align}
  N_{e,t} &\approx N \exp\left(- \frac{V_h}{2} \sum_{l=1}^L Q_{t,l}^2\right).
\end{align}

For a continuous recombination map, the recombination fraction between the
neutral allele and a selected site $d$ map distance away is given by Haldane's
map function, $r = (1-e^{-2d})/2$. Santiago and Caballero assume the focal
neutral site is in the middle of the chromosome, and integrate over the map
distance. Assuming that linkage is not too tight,

\begin{align}
  Q_\infty &= \frac{1}{1-r \kappa}
\end{align}

For our model, we consider the contribution of individual contiguous segments
of basepairs rather than the whole genome.

This summation can then in turn be
approximated by a integral; see \textcite{Santiago1998-bs} for more. Note that
another way to arrive at this equation is to take the product of reduction
factors $\nicefrac{N_e}{N}$ across selected sites, as in equation 4 of
\textcite{Hudson1995-xc}.

>>>>

\begin{align}
  \label{eq:qterm}
  S_1 Q_t &= S_1 + S_2' + S_3' + \ldots + S_t' \\
  Q_t &= 1 + \sum_{i=1}^t \left((1-r)\kappa\right)^i
\end{align}

which converges to,

\begin{align}
  \label{eq:Qinf}
  Q_\infty &= \frac{1}{1-(1-r) \kappa}.
\end{align}

This shows the long reach of heritable fitness factors in altering allele
frequency through the generations. In terms like $\E(S_1 S_2)$, a fraction
$S_1'$ of the heritable allele frequency change $S_2$ were \emph{previously}
built associations between the neutral allele and a fitness background from
$S_1$. As long as the direction of selection is the same, the expected change
of this fraction of associations formed in the first generation to later
generations 2 and 3 would be,

@@

First, we extend equation \eqref{eq:ne_t} to a polygenic system. Throughout, a
multiplicative fitness model is assumed (i.e. independent effects, no
epistasis), where the fitness of individual $i$ is the product of their fitness
contributions from each locus $l$, $w_{i,l}$, giving $w_i = \prod_{l=1}^n
w_{i,l}$. Then, assuming mean fitness is one and independence between sites,
the total fitness variation is multiplicative across the locus-level fitnesses,

\begin{align}
  V_h %&= \E_i\left( w_i^2 \right) - 1 \\
      %&= \prod_{l=1}^n (1 + \alpha_{l}^2 \V(g_l))  - 1  \\
      &= \prod_{l=1}^n (1 + V_{h,l})  - 1
\end{align}

Similarly, the $Q_t^2 V_h$ term in the single-locus model is multiplicative
under the polygenic model, as the change in site-specific fitness variances are
assumed independent,

\begin{align}
  Q_t^2 V_h &= \prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)  - 1
\end{align}

@@

\begin{align}
  N_{e,t} &\approx \frac{N}{V_h Q_t^2 + 1} \\
          &\approx \frac{N}{\prod_{l=1}^n \left(1 + \frac{V_{h,l} Q_{t,l}^2}{2}\right)} \\
          &\approx N \exp\left(- \sum_{l=1}^n \frac{V_{h,l} Q_{t,l}^2}{2}\right)
\end{align}

(c.f. \cite{Santiago1998-bs} equation 4b). If we write each site $l$'s
contribution to the variance as a deviation from the average contribution,
$V_{h,l}^2 = \nicefrac{1}{n} V_h + \varepsilon_l$, then

\begin{align}
  N_{e,t} &\approx N \exp\left(- \frac{1}{2}\sum_{l=1}^n \left(\frac{1}{n}V_{h} + \varepsilon_l\right)  Q_{t,l}^2\right) \\
          &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2+ \sum_{l=1}^n \varepsilon_l  Q_{t,l}^2\right)
\end{align}

Santiago and Caballero assume that over evolutionary replicates, as selected
sites are randomly distributed, there is not systematic covariance between the
fitness variation at a site and the decay of its association with the neutral
site, e.g. $\E(\varepsilon_l Q_{t,l}^2) = 0$, and thus ignore the second sum
term.

\begin{align}
      N_{e,t} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{t,l}^2\right) \\
\end{align}

@@

Our model differs from that of \textcite{Santiago1998-bs} in that we do not
integrate over the entire genome. In their 1998 model, Santiago and Caballero
consider the impact of both sites under selection on the same chromosome, as
well as sites on independently assorting chromosomes. We only consider the
contribution of a single segment under purifying selection, and the reduction
from additional segments accumulate multiplicatively (similar to our
implementation of the classic BGS model, XXX). Like their model, we imagine the
reduction experienced by a focal neutral site in the middle of the segment.
The implementation of our method differs slightly on though, see XXX.

Since $Q_{t,l}$ terms depend on $l$ only through the recombination fraction
between the neutral allele and the selected site, $r(l)$. Assuming a constant
per-basepair recombination rate of $r_\text{BP}$, $r(l) = r_\text{BP} l$ we can
approximate the sum with an integral for the asymptotic $Q_\infty$. The
reduction experienced in the middle of a chromosome is twice the reduction
experienced from each half,

\begin{align}
      N_{e,\infty} &\approx N \exp\left(- \frac{V_{h}}{2n} \sum_{l=1}^n Q_{\infty,l}^2\right) \\
              &\approx N \exp\left(- \frac{V_{h}}{L} \int_0^{\nicefrac{L}{2}} Q_{\infty}(r)^2 dr \right) \\
\end{align}

where $L = n r_\text{BP}$ is the total segment length in Morgans. Let us define
the total asymptotic effect of associations from all sites as $Q_\infty^2$.
Then,

\begin{align}
  Q_\infty^2 &= \frac{2}{L}\int_0^{\nicefrac{n r_\text{BP}}{2}} Q_{\infty}(x)^2 dx  \\
  &=  \frac{2}{L} \int_0^{\nicefrac{L}{2}} \left(\frac{1}{1 - (1-x)Z}\right)^2 dx  \\
  &= \frac{2}{(1-Z) (2-(2-L) Z)}
\end{align}

Santiago and Caballero further approximate this; we can derive their
approximation by setting $Z = 1-\kappa$ and doing a first-order Taylor series
expansion around $\kappa$ (since the loss in variance is presumed to be small).
Then,

\begin{align}
  Q_\infty^2 = \frac{2}{L (1-Z)} + \frac{2 (L-2)}{L^2}+\frac{2 \left(L^2-4 L+4\right) (1-Z)}{L^3} + \mathcal{O}(\kappa^2)
\end{align}

Santiago and Caballero
